Are populations like a circuit? Comparing isolation by resistance to a new coalescent-based method.

### 2019-04-20 review-responses.tex

\reply{
    We're not sure which possible study aims are being referred to,
    but one possible aim that is not possible currently with coalescence methods
    is to produce high-resolution maps (as we discuss, \revref).
    We made the statement that
    genetic data are more accurately modeled by coalescence time than resistance distance
    because while there is an extensive literature % going back fourty years
    describing the broad conditions under which coalescence is a good model 
    for the process that produces genomic data,
    we know of no such justification for resistance distance.
    Perhaps the key disconnect here is that we're thinking about generative models?
    We agree that since coalescence theory does not always apply,
    resistance distance on some network could in principle be closer to the truth
    than any coalescence times.
    However, we're unaware of this claim having been made.
    Furthermore, as we understand it,
    the point of landscape genetics methods using resistance distance
    is not merely to recapitulate observed genetic distances,
    but to provide an accurate depiction of past or present gene flow,
    i.e., movements of individuals across the landscape,
    In our paper we've tried to describe how inaccuracies of the generative model
    that underlies resistance distance
    can lead to misleading results.
    The only downside to coalescence time inference is computational tractability
    (discussed at \llname{ll:approx}):
    otherwise, one can be swapped out for the other.
    In fact, \emph{any} method that uses resistance distances could in principle be 
    done with coalescence times instead (e.g., to look at the effect of landscape features).
    This is discussed at \llname{ll:circuitscape}.
}

@@

\begin{point}{\revref}
    I think it  might be helpful if this point was made earlier and reiterated
\end{point}

\reply{
    Done, at \llname{ll:threepop}.
}

\reviewersection{5}

\begin{quote}
    This article focuses on the inference of gene flows between populations from
    the analysis of georeferenced genetic data. In particular, a new
    coalescent-based technique, which applies to the analysis of pairs of
    sampled individuals, is compared to an approach based on ``resistance
    distances'' which aim at approximating coalescent distances. This work is
    interesting because resistance distances are at the core of several popular
    techniques in landscape genetics. It is thus important to gain a good
    understanding of how good this phenomenological approximation is compared
    to more mechanistic techniques such as the structured coalescent. The
    article clarifies the connections between the two approaches in a
    convincing manner. It thereby fully deserves to be published in Molecular
    Ecology Resources. There are however a few points (see below) that I think
    should be improved should the authors see fit.
\end{quote}

@@

\begin{point}{}
    Regarding Equation 2, I would recommend giving more detail explaining where
    this comes from. Here is a suggestion:
    \begin{align*}
        H_{xz} &= dt + \sum_y Pr(X_{dt} = y | X_0 = x) H_{yz} \\
               &= dt + \sum_{y != x} Q_{xy} dt H_{yz} + (Q_{xx} dt + 1) H_{xz}
    \end{align*}
\end{point}

\reply{
    We wondered whether this would overload readers,
    but have now included an explanation along these lines.
    \revref
}

\begin{point}{}
    On a related note, it might be helpful to give more explanations about Equation 6. In
    fact, I am not quite sure I agree with it. Indeed, for $x \neq z$, we have
    \begin{align*}
        C_{xz} &= dt + {} \\
                & \qquad
             1/2 \sum_y Pr(X_{dt} = y | X_0 = x) C_{yz} + {} \\
                & \qquad
             1/2 \sum_y Pr(X_{dt} = y | X_0 = z) C_{xy}, \\
           &= dt + {} \\
                & \qquad
             1/2 \sum_{y \neq x} Q_{xy} dt C_{yz} + (Q_{xx} dt + 1) C_{xz} + \\
                & \qquad
             1/2 \sum_{y \neq x} Q_{zy} dt C_{xy} + (Q_{xx} dt + 1) C_{xz},
    \end{align*}
    where the 1/2 terms correspond to the probability of "selecting" the lineage found in x
    (resp. z) at time 0 to move to y after dt. The last equation above then gives
    \begin{align*}
        -2 = \sum_y Q_{xy} C_{yz} + Q_{zy} C_{xy}
    \end{align*}
    which differs from the one given in the ms. I have probably missed something in the derivation
    used by the authors. Giving a few more explanations here might help avoiding any confusion.
\end{point}

\reply{
    Besides substituting $G$ for $Q$,
    our equations only differ by the factor of 1/2.
    The model we work under is that each lineage moves as an independent copy
    of the Markov chain $X$, so that the total jump rate
    when the two chains are at $x$ and $z$ is $-G_{xx}-G_{zz}$.
    We think the factors of 1/2 are in error --
    a more clear way to see this would be to say that $X$ and $Z$ are the locations of the two lineages,
    that move independently according to $G$ until they coalesce,
    and so
    \begin{align*}
        C_{xz} &= dt + {} \\
                & \qquad
             \sum_u \P(X_{dt} = u | X_0 = x) C_{uz} + {} \\
                & \qquad
             \sum_v \P(Z_{dt} = v | Z_0 = z) C_{xv}, \\
           &= dt + {} \\
                & \qquad
             \sum_{u \neq x} G_{xu} dt C_{uz} + (G_{xx} dt + 1) C_{xz} + \\
                & \qquad
             \sum_{v \neq x} G_{zv} dt C_{xv} + (G_{xx} dt + 1) C_{xz} .
    \end{align*}
}
